### 2024-01-10 inst/scripts/papertest.Rmd
add paper like expe to try

adaptive traits and communities:
```{r}
km=5
ki=z-km
traits=initAdaptiveTraits(ki=ki,km=km,n=3) # this generate a table for 3 adaptive traits, with ki communities of full hunter-gathere and km of farmers

G=5 #limits/numer of row & column of the world grid
coordinates=random2Dgrid(K=ki+km,Gx=G) #random coordinated on the grid

@@

```{r runburnin,warning=FALSE,message=FALSE,results='hold',cache=TRUE}
tstep=500
burnin=modelVector(
   K=ki, m=1, b=0.2, r=0, rho=0, d=0.01, maturity=18, endrepro=65, 
   population=population, 
   comus=initcomus, 
   tstep=tstep, 
   tp=initNeutralTraitsPathways(z = z) , #we recreate a temporary neutral trait table with no neutral transmission
   age.threshold=20, 
   out=c("finalpop","finalcomus"),   #we only want the final population
   logging=c("time"),ma=1,traitsid=paste0("t",1:z),F_Th = 100
)
population=burnin$population
initcomus$size=burnin$finalcomus$size
```

```{r setuppopulation}
traitsid=paste0("t",1:z)
population[,traitsid]=0
for(i in 1:z){population[population[,"community"]==i,traitsid[i]]=1}
knitr::kable(aggregate(population[,traitsid],by=list(population[,"community"]),sum))
```

We can also reset ids to start at one and have more readable results:

```{r resetids}
population[population[,"cid"]!= -1 ,"cid"] = as.numeric(as.factor(as.character(population[population[,"cid"]!= -1 ,"cid"])))
population[,"fid"] = as.numeric(as.factor(as.character(population[ ,"fid"])))
ids = as.numeric(as.factor(as.character(population[ ,"id"])))
names(ids)=population[,"id"]
population[,"id"] = as.numeric(as.factor(as.character(population[ ,"id"])))
population[,"partner"]=as.numeric(ids[as.character(population[,"partner"])])
population[is.na(population[,"partner"]),"partner"]=-1
```

```{r runpropersimu}
tstep=1000
a=Sys.time()
initcomus$occupation[,]=1 #cheat and prohibit the empty space to be colonised
simuTest=modelVector(
    K=ki, m=1, b=0.2, r=0.01, rho=0, d=0.01, 
    maturity=18, endrepro=65, 
    population=population, 
    comus=initcomus, 
    tstep=tstep, 
    tp=neutraltraitsParam,
    age.threshold=20, 
    out=c("finalpop","finalcomus","traitsumary","popsize"), #this time we keep track of more things
    logging=c("time","done","visu"),
    beta=-1,
    ma=1,traitsid=paste0("t",1:z),F_Th = 150
)
print(Sys.time()-a)
```

@@

```{r followpopsize}
plot(simuTest$popsize,type="l",ylim=c(300,650),lwd=5,xlab="time",ylab="number of individual")
```

@@

```{r runsimu-pre,warnings=FALSE,results=FALSE,results='hold'}

@@

```{r,eval=F}
tstep=500
repli=50
neutraltraitsParam$s=c(.5,.5,.5,.5)
rhodep=lapply(c(0,1),function(rho){
           neutraltraitsParam$post[,]=0
           neutraltraitsParam$pre[,]=0
           neutraltraitsParam$post[,"i"]=1
           allinlaw=replicate(repli,
                              {
                                  quickV=modelVector(K=K, m=1, b=0.2, r=0.01, rho=rho, d=0.01, maturity=18, endrepro=65, population=population, comus=initcomus, tstep=tstep, tp=neutraltraitsParam,age.threshold=20, out=c("finalpop","finalcomus","popsize"),logging=c("time"),ma=1,traitsid=paste0("t",1:z),F_Th = 150,vidfile=NULL)
                                  pop=quickV$population
                                  sapply(1:z,function(i)apply(pop[pop[,"community"]==i,traitsid],2,sum))
                              },simplify=F)

>>>>

adaptive traits and communities:
```{r}
km=1
ki=4
traits=initAdaptiveTraits(ki=ki,km=km,n=3) # this generate a table for 3 adaptive traits, with ki communities of full hunter-gathere and km of farmers
traits[1:ki,]=0
traits[5,]=1

G=5 #limits/numer of row & column of the world grid
coordinates=random2Dgrid(K=ki+km,Gx=G) #random coordinated on the grid
farmers=expand.grid(3:5, 3:5)
coordinates[1:ki,]=as.matrix(farmers[sample(1:nrow(farmers),ki),])
coordinates[5,]=c(1,1)

@@

```{r runburnin,warning=FALSE,message=FALSE,results='hold',cache=TRUE}
tstep=500
tpcomu=initcomus
tpcomu$occupation[,]=1
burnin=modelVector(
   K=ki, m=1, b=0.2, r=0, rho=0, d=0.01, maturity=18, endrepro=65, 
   population=population, 
   comus=tpcomu, 
   tstep=tstep, 
   tp=initNeutralTraitsPathways(z = z) , #we recreate a temporary neutral trait table with no neutral transmission
   age.threshold=20, 
   out=c("finalpop","finalcomus"),   #we only want the final population
   beta=-1,
   logging=c("time"),ma=1,traitsid=paste0("t",1:z),F_Th = 50
)
population=resetIds(burnin$population)
initcomus$size=burnin$finalcomus$size
```

```{r setuppopulation}
population[,traitsid]=0
population[population[,"community"]==5,traitsid]=1
knitr::kable(aggregate(population[,traitsid],by=list(population[,"community"]),sum))

```{r runpropersimu}
tstep=1000
a=Sys.time()
simuTest=modelVector(
    K=ki, m=1, b=0.2, r=0.01, rho=0, d=0.01, 
    maturity=18, endrepro=65, 
    population=population, 
    comus=initcomus, 
    tstep=tstep, 
    tp=neutraltraitsParam,
    age.threshold=20, 
    out=c("finalpop","finalcomus","traitsumary","popsize"), #this time we keep track of more things
    logging=c("time","done","visu"),
    beta=-1,
    ma=1,traitsid=paste0("t",1:z),F_Th = 100
)
print(Sys.time()-a)

@@

```{r followpopsize}
plot(simuTest$popsize,type="l",ylim=c(300,2500),lwd=5,xlab="time",ylab="number of individual")
```

@@

```{r track-traits,warnings=FALSE,results=FALSE,results='hold'}

traitsfreq=simuTest$traitsumary/simuTest$popsize

pathways= c("in-law","horiz before","horiz after")
plot(1,1,ylim=c(0,1),xlim=c(0,1000),type="n",xlab="time",ylab="freq")
sapply(1:24,function(tr)lines(traitsfreq[,tr],col=tr))
par(xpd=NA)
legend(x=-9,y=1.5,lwd=4,col=1:4,legend=paste0("t",1:4),cex=1,bty="n")

@@

```{r,eval=F}
tstep=20000
repli=50
neutraltraitsParam$s=c(.5,.5,.5,.5)
rhodep=lapply(c(0,1),function(rho){
           neutraltraitsParam$post[,]=0
           neutraltraitsParam$pre[,]=0
           neutraltraitsParam$post[,"i"]=1
           allinlaw=replicate(repli,
                              {
                                  quickV=modelVector(K=K, m=1, b=0.2, r=0.01, rho=rho, d=0.01, maturity=18, endrepro=65, population=population, comus=initcomus, tstep=tstep, tp=neutraltraitsParam,age.threshold=20, out=c("finalpop","finalcomus","popsize"),logging=c("time"),ma=1,traitsid=paste0("t",1:z),F_Th = 150,vidfile=NULL)
                                  pop=quickV$population
                                  sapply(1:z,function(i)apply(pop[pop[,"community"]==i,traitsid],2,sum))
                              },simplify=F)
